\plaintitle

### 2017-01-08 drafts/labor/piecework_lit.tex

>>>>

% \onlyinsubfile{\clearpage}
\subsubsection{\ali{should I drop this? Railroad and other }Industrial Workers}
\topic{Piecework might be most familiar to the HCI researcher at the assembly line, which largely defined manufacturing through the \nth{20} century.}
In the railroad industry and especially on the assembly line
we see many of the mechanisms of work and worker management that later made their way into crowd work.
Furthermore, it's from here the piecework literature draws most upon to find limits speaking to complexity and task granularity.
We'll expose these facets of piecework very superficially in the context of industrial work
to better speak to these research questions in on--demand labor later.

\topic{Railway companies adopted piecework regimes in the early \nth{20} century;
what followed was a flourishing in the management practices with pieceworkers,
as railway companies struggled to find effective ways
to motivate and evaluate this skilled workforce of engineers.}
\citeauthor{10.2307/23702539} takes up a case study of the Santa Fe Railway,
finding that they employed ``efficiency experts'' to develop a ``standard time''
to determine pay for each task at the company informed by
``thousands of individual operations''; % ~\cite{10.2307/23702539}.
\citeauthor{10.2307/23702539} goes on to list
some of the roles required to facilitate piecework
in the early \nth{20} century
--- among them, ``piecework clerks, inspectors, and `experts'~''~\cite{10.2307/23702539}.

\topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
found its way into the war effort during World War II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a predominantly female workforce that had neither
the formal training nor
the years of apprenticeship experience
that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
% --- such as riveting component to another ---
and assigned to that task. %, decomposed from a more comprehensive task.
One might reflect on the observation that ``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
was a pieceworker~\cite{davies2014origins}.

% \msb{even further backwards??? I would have expected this section to take a time slice and tell me about one group of people in detail}
% \msb{this content might go above in the paragraph about the heydey of piecework in the piecework review}
\topic{The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.}
\citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
They attribute this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.
